---
title: "Natural Resource Curse EITI"
author: "Monique Bennett"
date: "26/01/2021"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(wbstats)
library(tidyverse)
library(democracyData)
library(ggpubr)
library(ggrepel)
library(extrafont)
library(tikzDevice)
loadfonts(device = "win", quiet = T)
```

## How does the resource curse work? 

```{r data, include=FALSE}
wgi_all <- download_wgi_voice_and_accountability()

indicators <- c(gdp_pc = "NY.GDP.PCAP.CD",
                gdp = "NY.GDP.MKTP.CD",
                resource_rents = "NY.GDP.TOTL.RT.ZS",
                fuel_exports =  "TX.VAL.FUEL.ZS.UN",
                imports = 'NE.IMP.GNFS.CD',
                exports = "NE.EXP.GNFS.CD","CC.EST", "GE.EST","RQ.EST", 
                "PV.EST", "RL.EST", "VA.EST")

africa_iso <- c("DZA", "AGO", "BEN", "BWA", "BFA", "BDI", "CV", "CMR", "CAF", "TCD", "COM", "COD",
                "COG", "CIV", "DJI", "EGY", "ERI", "SWZ", "ETH", "GAB", "GMB", "GHA", "GIN", "GNB",
                "KEN", "LSO", "LBR", "LBY", "MDG", "MWI", "MLI", "MRT", "MUS", "MAR", "MOZ", "NAM",
                "NER", "NGA", "RWA", "STP", "SEN", "SYC", "SLE", "SOM", "ZAF", "SSD", "SDN", "TZA",
                "TGO", "TUN", "UGA", "ZMB", "ZWE")

wb <- wb_data(country = africa_iso, indicator = indicators, start_date = 2018, 
              end_date = 2018, return_wide = TRUE)

wgi_africa <- wgi_all %>% 
  filter(wb_code %in% africa_iso, year == 2018)

#allcountries <- read_csv("data/countryincomelist.csv")

#developing <- subset(allcountries[, c("country_name", "iso3c", "region", 
#                                      "income_class")], income_class != "High income")

HDI <- readxl::read_excel("C:/Users/MoniqueBennett/OneDrive - GGA Good Governance Africa NPC/Documents/resource_curse_africa/data/HDI_data.xlsx")

HDI$iso3c <- countrycode::countrycode(HDI$Country, origin = "country.name", destination = "iso3c")

HDI_africa <- HDI %>% 
  filter(iso3c %in% africa_iso)

## LIED

lied <- readxl::read_excel("C:/Users/MoniqueBennett/OneDrive - GGA Good Governance Africa NPC/Documents/resource_curse_africa/data/lied_v5.2.xls")

lied_2019 <- subset(lied[, c("countryn", "year", "lexical_index")], year == 2019)

lied_2019$iso3c <- countrycode::countrycode(lied_2019$countryn, origin = "country.name", destination = "iso3c")

lied_africa <- lied_2019 %>% 
  filter(iso3c %in% africa_iso)

lied_africa$dummy <- NA

lied_africa$dummy[lied_africa$lexical_index == 0] <- 0

lied_africa$dummy[lied_africa$lexical_index == 1] <- 0

lied_africa$dummy[lied_africa$lexical_index == 2] <- 0

lied_africa$dummy[lied_africa$lexical_index == 3] <- 0

lied_africa$dummy[lied_africa$lexical_index == 6] <- 1

lied_africa$dummy2 <- as.numeric(lied_africa$dummy)

lied_africa$dummy2[lied_africa$dummy2 == 0] <- "Non-democratic"
lied_africa$dummy2[lied_africa$dummy2 == 1] <- "Democratic"

lied_africa$Regime <- as.factor(lied_africa$dummy2)

wb <- merge(wb, lied_africa, by = 'iso3c')

```

```{r}
#wb %>%
 # select(country, CC.EST, GE.EST, exports, imports, gdp_pc, resource_rents, VA.EST) %>% 
 # knitr::kable()

summary(wb)
```

### Top 10 Resource Rents in Africa

Total natural resources rents are the sum of oil rents, natural gas rents, coal rents (hard and soft), mineral rents, and forest rents.

```{r}
wb %>% 
  select(country, resource_rents, GE.EST, VA.EST) %>% 
  slice_max(resource_rents, n = 10) %>% 
  knitr::kable()
```

### Bottom 10 Resource Rents in Africa

```{r}
wb %>% 
  select(country, resource_rents, GE.EST, VA.EST) %>%
  slice_min(resource_rents, n = 10) %>% 
  knitr::kable()
```

### Top 10 Fuel export countries in Africa

```{r}
wb %>% 
  select(country, fuel_exports, GE.EST, VA.EST) %>%
  slice_max(fuel_exports, n = 10) %>% 
  knitr::kable()
```

## Trade calculations

```{r}
wb$trade <- (wb$imports + wb$exports) / wb$gdp

wb$trade_resources <- wb$fuel_exports / wb$imports
```
 
### Correlation 

```{r}
cor.test(wb$resource_rents, wb$gdp_pc, method=c("pearson", "kendall", "spearman"))

cor.test(wb$resource_rents, wb$CC.EST, method=c("pearson", "kendall", "spearman"))

cor.test(wb$resource_rents, wb$GE.EST, method=c("pearson", "kendall", "spearman"))

cor.test(wb$fuel_exports, wb$resource_rents, method=c("pearson", "kendall", "spearman"))
```

### Correlation Plots

Resource Rents vs GDPpc
```{r warning=F}
countries2 <-  c('Angola', "Botswana", 'Comoros', "Congo, Dem. Rep.", 'Eswatini', 'Lesotho', 'Madagascar', 'Malawi', 'Mauritius', 'Mozambique', 'Namibia', 'Seychelles', 'South Africa', 'Tanzania', 'Zambia', 'Zimbabwe')

countries <- wb %>% 
  filter(country %in% countries2) %>% 
  select(country, resource_rents, CC.EST, gdp_pc, VA.EST)


ggscatter(wb, x = "resource_rents", y = "gdp_pc", 
          add = "reg.line", conf.int = TRUE, palette = "jco",
          cor.method = "pearson", cor.coef = T, cor.coef.coord = c(-2.7, 2.5),
          shape = "Regime",
          xscale = "log10", yscale = "log10", 
          font.family = "Proxima Nova Rg",
          label.select = countries2, label = "country", repel = T,
          title = "What does wealth in resources mean for economic prosperity in Africa?",
          xlab = "Total natural resources rents (log scale)", ylab = "GDP per capita (log scale)") +
  scale_shape_manual(values = c(19, 1)) +
  labs(caption = '*Total natural resources rents are the sum of oil, natural gas, coal, mineral and forest rents.
Source: World Bank Development Indicators and Lexical Index of Electoral Democracy 26 January 2020. 
Graphic: Monique Bennett at Good Governance Africa',
subtitle = "Latest data on total natural resource rents and its relationship with GDP per capita remains negative.") +
  theme(plot.caption = element_text(hjust = 0, size = 9), plot.caption.position = "plot", plot.subtitle = element_text(size = 9),
        plot.title = element_text(face = "bold", size = 12), legend.position = "bottom",
          legend.box = "vertical")

ggsave("resource_rents_gdppc.png")
```

Resource Rents vs Government Effectiveness
```{r warning=F}
ggscatter(wb, x = "resource_rents", y = "GE.EST", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xscale = "log10",
          xlab = "Total natural resources rents", ylab = "Government Effectiveness WGI")
```

Resource Rents vs Control of Corruption
```{r warning=F}
options(ggrepel.max.overlaps = Inf)

ggscatter(wb, x = "resource_rents", y = "CC.EST", 
          add = "reg.line", 
          conf.int = TRUE, 
          palette = "jco",
          cor.method = "pearson", cor.coef = T, cor.coef.coord = c(-2.7, -2.4),
          shape = "Regime",
          xscale = "log10", 
          font.family = "Proxima Nova Rg",
          #label.select = countries2, label = "country", repel = T,
          title = "Poor control over corruption is high among African states with high resource rents.",
          xlab = "Total natural resources rents (log scale)", ylab = "Control of Corruption") +
  scale_shape_manual(values = c(19, 1)) +
  labs(caption = '*Control of corruption captures perceptions of the extent to which public power is exercised for private gain, 
including both petty and grand forms of corruption, as well as "capture" of the state by elites and private interests.
Source: World Bank Development & Governance Indicators and Lexical Index of Electoral Democracy 26 January 2020. 
Graphic: Monique Bennett at Good Governance Africa',
subtitle = "African countries with poor corruption scores (less than 1) have higher resource rents, which suggests 
a negative relationship. Regime type is shown using point shape.") +
  geom_text_repel(data = countries, aes(resource_rents, CC.EST, label = country), size = 4, color = "black", force = 4,
                  box.padding = 0.5) +
  theme(plot.caption = element_text(hjust = 0, size = 10), plot.caption.position = "plot", 
        plot.subtitle = element_text(size = 10),
        plot.title = element_text(face = "bold", size = 12), legend.position = "bottom",
          legend.box = "vertical") +
  expand_limits(y = c(-2.5, 2.5)) 


ggsave("corruption_resource_rents.png", width = 8, height = 8)
```

Resource Rents vs Voice & Accountability
```{r warning=F}
ggscatter(wb, x = "resource_rents", y = "VA.EST", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xscale = "log10",
          xlab = "Total natural resources rents", ylab = "Voice & Accountability WGI")
```

Trade of Resources vs Voice & Accountability
```{r}
ggscatter(wb, x = "trade_resources", y = "VA.EST", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Trade resources", ylab = "Voice & Accountability WGI")
```

Trade of Resources vs Control of Corruption 
```{r}
ggscatter(wb, x = "trade_resources", y = "CC.EST", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Trade resources", ylab = "Voice & Accountability WGI")
```

### Human Development Index and Resource Rents

```{r}
HDI_wb <- merge(wb, HDI, by.all = "ïso3c")
```

### Correlation 

```{r}
cor.test(HDI_wb$resource_rents, HDI_wb$HDI, method=c("pearson", "kendall", "spearman"))

cor.test(HDI_wb$fuel_exports, HDI_wb$HDI, method=c("pearson", "kendall", "spearman"))

cor.test(HDI_wb$trade, HDI_wb$HDI, method=c("pearson", "kendall", "spearman"))

cor.test(HDI_wb$VA.EST, HDI_wb$HDI, method=c("pearson", "kendall", "spearman"))
```
### Correlation Plots

Resource Rents vs HDI
```{r warning=F}
ggscatter(HDI_wb, x = "resource_rents", y = "HDI", 
          add = "reg.line", conf.int = TRUE, palette = "jco",
          cor.coef = TRUE, cor.method = "pearson",
          shape = "Regime",
          xscale = "log10", yscale = "log10", 
          font.family = "Proxima Nova Rg",
          label = "country", repel = T,
          title = "What does wealth in resources mean for human development in Africa?",
          xlab = "Total natural resources rents (log scale)",  ylab = "Human Development Index") +
  scale_shape_manual(values = c(19, 1)) +
  labs(caption = '*Total natural resources rents are the sum of oil, natural gas, coal, mineral and forest rents.
Source: World Bank Development Indicators and Human Development Index 26 January 2020. 
Graphic: Monique Bennett at Good Governance Africa',
subtitle = "Latest data on total natural resource rents and its relationship with human development is negative.") +
  theme(plot.caption = element_text(hjust = 0, size = 8), plot.caption.position = "plot", plot.subtitle = element_text(size = 9),
        plot.title = element_text(face = "bold", size = 11), legend.position = "bottom",
          legend.box = "vertical")

ggsave("humandevelopment_resourcerents.png")
```

Fuel exports vs HDI
```{r warning=F}
ggscatter(HDI_wb, x = "fuel_exports", y = "HDI", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xscale = "log10",
          xlab = "Fuel exports", ylab = "Human Development Index")
```

Trade resources vs HDI
```{r warning=F}
ggscatter(HDI_wb, x = "trade_resources", y = "HDI", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Trade resources", ylab = "Human Development Index")
```