---
title: "Natural Resource Curse EITI"
author: "Monique Bennett"
date: "26/01/2021"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(wbstats)
library(tidyverse)
library(democracyData)
library(ggpubr)
```

## How does the resource curse work? 

```{r data, include=FALSE}
wgi_all <- download_wgi_voice_and_accountability()

indicators <- c(gdp_pc = "NY.GDP.PCAP.CD",
                gdp = "NY.GDP.MKTP.CD",
                resource_rents = "NY.GDP.TOTL.RT.ZS",
                fuel_exports =  "TX.VAL.FUEL.ZS.UN",
                imports = 'NE.IMP.GNFS.CD',
                exports = "NE.EXP.GNFS.CD","CC.EST", "GE.EST","RQ.EST", 
                "PV.EST", "RL.EST", "VA.EST")

africa_iso <- c("DZA", "AGO", "BEN", "BWA", "BFA", "BDI", "CV", "CMR", "CAF", "TCD", "COM", "COD",
                "COG", "CIV", "DJI", "EGY", "ERI", "SWZ", "ETH", "GAB", "GMB", "GHA", "GIN", "GNB",
                "KEN", "LSO", "LBR", "LBY", "MDG", "MWI", "MLI", "MRT", "MUS", "MAR", "MOZ", "NAM",
                "NER", "NGA", "RWA", "STP", "SEN", "SYC", "SLE", "SOM", "ZAF", "SSD", "SDN", "TZA",
                "TGO", "TUN", "UGA", "ZMB", "ZWE")

wb <- wb_data(country = africa_iso, indicator = indicators, start_date = 2018, 
              end_date = 2018, return_wide = TRUE)

wgi_africa <- wgi_all %>% 
  filter(wb_code %in% africa_iso, year == 2018)

#allcountries <- read_csv("data/countryincomelist.csv")

#developing <- subset(allcountries[, c("country_name", "iso3c", "region", 
#                                      "income_class")], income_class != "High income")

HDI <- readxl::read_excel("C:/Users/MoniqueBennett/OneDrive - GGA Good Governance Africa NPC/Documents/resource_curse_africa/data/HDI_data.xlsx")

HDI$iso3c <- countrycode::countrycode(HDI$Country, origin = "country.name", destination = "iso3c")

HDI_africa <- HDI %>% 
  filter(iso3c %in% africa_iso)

```

```{r}
#wb %>%
 # select(country, CC.EST, GE.EST, exports, imports, gdp_pc, resource_rents, VA.EST) %>% 
 # knitr::kable()

summary(wb)
```

### Top 10 Resource Rents in Africa

Total natural resources rents are the sum of oil rents, natural gas rents, coal rents (hard and soft), mineral rents, and forest rents.

```{r}
wb %>% 
  select(country, resource_rents, GE.EST, VA.EST) %>% 
  slice_max(resource_rents, n = 10) %>% 
  knitr::kable()
```

### Bottom 10 Resource Rents in Africa

```{r}
wb %>% 
  select(country, resource_rents, GE.EST, VA.EST) %>%
  slice_min(resource_rents, n = 10) %>% 
  knitr::kable()
```

### Top 10 Fuel export countries in Africa

```{r}
wb %>% 
  select(country, fuel_exports, GE.EST, VA.EST) %>%
  slice_max(fuel_exports, n = 10) %>% 
  knitr::kable()
```

## Trade calculations

```{r}
wb$trade <- (wb$imports + wb$exports) / wb$gdp

wb$trade_resources <- wb$fuel_exports / wb$imports
```
 
### Correlation 

```{r}
cor.test(wb$resource_rents, wb$gdp_pc, method=c("pearson", "kendall", "spearman"))

cor.test(wb$resource_rents, wb$CC.EST, method=c("pearson", "kendall", "spearman"))

cor.test(wb$resource_rents, wb$GE.EST, method=c("pearson", "kendall", "spearman"))

cor.test(wb$fuel_exports, wb$resource_rents, method=c("pearson", "kendall", "spearman"))
```

### Correlation Plots

Resource Rents vs GDPpc
```{r warning=F}
ggscatter(wb, x = "resource_rents", y = "gdp_pc", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xscale = "log10", yscale = "log10",
          xlab = "Total natural resources rents (log)", ylab = "GDP per capita (log)")
```

Resource Rents vs Government Effectiveness
```{r warning=F}
ggscatter(wb, x = "resource_rents", y = "GE.EST", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xscale = "log10",
          xlab = "Total natural resources rents", ylab = "Government Effectiveness WGI")
```

Resource Rents vs Control of Corruption
```{r warning=F}
ggscatter(wb, x = "resource_rents", y = "CC.EST", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xscale = "log10",
          xlab = "Total natural resources rents", ylab = "Control of Corruption WGI")

```

Resource Rents vs Voice & Accountability
```{r warning=F}
ggscatter(wb, x = "resource_rents", y = "VA.EST", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xscale = "log10",
          xlab = "Total natural resources rents", ylab = "Voice & Accountability WGI")
```

Trade of Resources vs Voice & Accountability
```{r}
ggscatter(wb, x = "trade_resources", y = "VA.EST", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Trade resources", ylab = "Voice & Accountability WGI")
```

Trade of Resources vs Control of Corruption 
```{r}
ggscatter(wb, x = "trade_resources", y = "CC.EST", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Trade resources", ylab = "Voice & Accountability WGI")
```

### Human Development Index and Resource Rents

```{r}
HDI_wb <- merge(wb, HDI, by.all = "ïso3c")
```

### Correlation 

```{r}
cor.test(HDI_wb$resource_rents, HDI_wb$HDI, method=c("pearson", "kendall", "spearman"))

cor.test(HDI_wb$fuel_exports, HDI_wb$HDI, method=c("pearson", "kendall", "spearman"))

cor.test(HDI_wb$trade, HDI_wb$HDI, method=c("pearson", "kendall", "spearman"))

cor.test(HDI_wb$VA.EST, HDI_wb$HDI, method=c("pearson", "kendall", "spearman"))
```
### Correlation Plots

Resource Rents vs HDI
```{r warning=F}
ggscatter(HDI_wb, x = "resource_rents", y = "HDI", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xscale = "log10",
          xlab = "Total natural resources rents", ylab = "Human Development Index")
```

Fuel exports vs HDI
```{r warning=F}
ggscatter(HDI_wb, x = "fuel_exports", y = "HDI", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xscale = "log10",
          xlab = "Fuel exports", ylab = "Human Development Index")
```

Trade resources vs HDI
```{r warning=F}
ggscatter(HDI_wb, x = "trade_resources", y = "HDI", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Trade resources", ylab = "Human Development Index")
```